---
title: "San Francisco Crime Heatmap"
author: "Nathaniel Reed"
date: "January 3, 2016"
output: html_document
---

Here I show how to present the San Francisco crime data on a heatmap that shows the crime numbers by police district.

We'll use the training set:

```{r}
unzip("train.csv.zip")
train <- read.csv("train.csv", 
                  colClasses=c("Date", "factor", "character", "factor", "factor", "factor", "character", "numeric", "numeric"))
summary(train)
```

Since this is spatial data, it would be nice if we could plot this on a map. Fortunately, the city of San Francisco makes this freely available:

```{r}
download.file("http://apps.sfgov.org/datafiles/view.php?file=sfgis/sfpd_districts.zip", method="curl", destfile="sfpd_districts.zip")
unzip("sfpd_districts.zip", exdir="sfpd_districts")
```

We use the rgdal package to load the shapefile into a data frame 'sfpd':

```{r}
library(rgdal)
sfpd <- readOGR(dsn="sfpd_districts", layer="sfpd_districts")
```

Next we aggregate crime data by police district and join it with the spatial data:

```{r}
library(dplyr)
# Rollup crime stats by police dept
by_pd <- group_by(train, PdDistrict)
by_pd <- by_pd %>% mutate(count = n())
by_pd_summary <- summarize(by_pd, num_crimes=n())

# Join with lnd by DISTRICT
by_pd_summary <- rename(by_pd_summary, DISTRICT=PdDistrict)
sfpd@data <- left_join(sfpd@data, by_pd_summary)
```

Now we can plot this on a heatmap:

```{r}
library(tmap)
qtm(sfpd, "num_crimes", fill.title="Number of Crimes") +
  tm_text("DISTRICT", size="AREA", root=5)
```

